---
title: "Assignment 2"
author: "Sara Arman"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load libraries, message=FALSE}}
library(tidycensus)
library(tidyverse)
library(ggplot2)
library(ggthemes)
library(viridis)
```

### For my first step, I am loading my data set from Assignment 1. 

This data set includes the following variables:
1. Sex, "SEX", categorical 
2. Wages or salary income past 12 months, "WAGP", continuous
3. Public assistance income past 12 months, "PAP", continuous
4. Educational attainment, "SCHL", categorical 
5. Citizenship status, CIT, categorical 
6. Age, AGEP, continuous

I have filtered my data to only include sex=2 (female) and wages and public assistance income greater than -1. This has narrowed my number of observations down from 70,131 to 31,262. 
```{rload and attach data, message=FALSE}
person_data <-read_csv('SaraArman-DataSet1_Final.csv') %>%
  filter(WAGP> -1 & PAP> -1 & SEX== 2)
```
## Continous Variables 

### My first continous variable is Wages over the past 12 months (WAGP) so I wil find the:
1. mean
2. standard deviation 
3. 95% CI for Population Mean 
4. Inter Quartile Range 
5. And then I will create a histogram.

```{r quick summary statistics}
WAGP_summary <- summary(person_data$WAGP)
WAGP_summary
```
 
 ### Mean
 
```{r}
WAGP_summary["Mean"]
```

### Standard Deviation

```{r}
sd(person_data$WAGP)
```


### 95% CI for Population mean 
```{r}
conf_int <- t.test(person_data$WAGP)
conf_int
```

### IQR (IQR = Q3-Q1 = 47000)

```{r}
quantile(person_data$WAGP)
```


### Histogram 

This histogram was not surprising, 
```{r}
options(scipen = 999)
ggplot(person_data, aes(x = WAGP)) +
  geom_histogram(color = "black" , bins = 20 )
scale_x_continuous(name = "Wages and Income Over the Past 12 Months")
                     breaks = breaks <- seq(0, 1000, by = 1000)
```

### My second continous variable is Public Assistance Income Recieved over past 12 months, PAP

```{r quick summary statistics}
PAP_summary <- summary(person_data$PAP)
PAP_summary
```

### Mean
 
```{r}
PAP_summary["Mean"]
```

### Standard Deviation

```{r}
sd(person_data$PAP)
```
### 95% CI 

```{r}
conf_int <- t.test(person_data$PAP)
conf_int
```

### IQR (IQR = Q3-Q1?)

```{r}
quantile(person_data$PAP)
```

### Histogram 
```{r}
ggplot(person_data,
       aes(x = PAP)) + 
geom_histogram(aes(fill = ..count..), bins = 8, color = "black", weight = 3) +
scale_x_continuous(name = "Public Assistance Income",
                   breaks = seq(0, 10000, by = 5000)) +
scale_y_continuous(name = "Count",
                  breaks = seq(0, 30000, by = 5000)) + 
labs(title = 'Public Assistance Income Received Over the Past 12 Months')+
 theme_bw()
coord_flip()
```


### My third continous variable is age

```{r quick summary statistics}
AGEP_summary <- summary(person_data$AGEP)
AGEP_summary
```

### Mean
 
```{r}
AGEP_summary["Mean"]
```
### Standard Deviation

```{r}
sd(person_data$AGEP)
```
### 95% CI 

```{r}
conf_int <- t.test(person_data$AGEP)
conf_int
```
### Histogram 
```{r}
ggplot(person_data, aes(x = AGEP)) +
  geom_histogram()


ggplot(person_data,
       aes(x = AGEP)) + 
geom_histogram(aes(fill = ..count..), bins = 10, color = "black", weight = 5) +
scale_x_continuous(name = "Age",
                   breaks = seq(0, 95, by = 10)) +
scale_y_continuous(name = "Count",
                  breaks = seq(0, 30000, by = 5000)) + 
labs(title = 'Age')+
 theme_bw()
coord_flip()
```


## Categorical Variables
For each categorical variable in my data set (SEX, SCHL, CIT), I will calculate the 95-percent confidence interval for the proportion of the population in each category.

### My first categorical variable is Sex

```{r}
unique(person_data$SEX_label)
```

```{r}
table(person_data$SEX_label)
```

```{r}
table(person_data$SEX_label) / sum(table(person_data$SEX_label))
```

```{r}
mean(person_data$SEX_label == "Female")
```

```{r}
conf_int <- t.test(person_data$SEX_label == "Female")
conf_int <- t.test(person_data$SEX_label == "Male")
```

### My second categorical variable is Educational Attainment (SCHL) 
I am finding the confidence interval for each category within educational attainment.

```{r}
conf_int <- t.test (person_data$SCHL == "1 or more years of college credit, no degree")
conf_int <- t.test (person_data$SCHL == "12th grade - no diploma")
conf_int <- t.test (person_data$SCHL == "Associate's degree")
conf_int <- t.test (person_data$SCHL == "Bachelor's degree")
conf_int <- t.test (person_data$SCHL == "Doctorate degree")
conf_int <- t.test (person_data$SCHL == "GED or alternative credential")
conf_int <- t.test (person_data$SCHL == "Grade 1")
conf_int <- t.test (person_data$SCHL == "Grade 10")
conf_int <- t.test (person_data$SCHL == "Grade 11")
conf_int <- t.test (person_data$SCHL == "Grade 2")
conf_int <- t.test (person_data$SCHL == "Grade 3")
conf_int <- t.test (person_data$SCHL == "Grade 4")
conf_int <- t.test (person_data$SCHL == "Grade 5")
conf_int <- t.test (person_data$SCHL == "Grade 6")
conf_int <- t.test (person_data$SCHL == "Grade 7")
conf_int <- t.test (person_data$SCHL == "Grade 8")
conf_int <- t.test (person_data$SCHL == "Grade 9")
conf_int <- t.test (person_data$SCHL == "Kindergarten")
conf_int <- t.test (person_data$SCHL == "Master's degree")
conf_int <- t.test (person_data$SCHL == "No schooling completed")
conf_int <- t.test (person_data$SCHL == "Nursery school, preschool")
conf_int <- t.test (person_data$SCHL == "Professional degree beyond a bachelor's degree")
conf_int <- t.test (person_data$SCHL == " Regular high school diploma")
conf_int <- t.test (person_data$SCHL == "Some college, but less than 1 year")
```

### My third categorical variable is Citizenship status , CIT
```{r}
unique(person_data$CIT_label)
```

```{r}
table(person_data$CIT_label)
```

```{r}
table(person_data$CIT_label) / sum(table(person_data$CIT_label))
```

```{r}
conf_int <- t.test (person_data$CIT == "Born abroad of American parent(s)")
conf_int <- t.test (person_data$CIT == "Born in Puerto Rico, Guam, the U.S. Virgin Islands, or the Northern Marianas ")
conf_int <- t.test (person_data$CIT == "Born in the U.S.")
conf_int <- t.test (person_data$CIT == "Not a citizen of the U.S. ")
conf_int <- t.test (person_data$CIT == "U.S. citizen by naturalization ")
```