---
title: "Assignment 2"
author: "Sara Arman"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load libraries, message=FALSE}
library(tidycensus)
library(tidyverse)
library(ggplot2)
library(ggthemes)
library(viridis)
```

### For my first step, I am loading my data set from Assignment 1. 

This data set includes the following variables:
1. Sex, "SEX", categorical 
2. Wages or salary income past 12 months, "WAGP", continuous
3. Public assistance income past 12 months, "PAP", continuous
4. Educational attainment, "SCHL", categorical 
5. Citizenship status, CIT, categorical 
6. Age, AGEP, continuous

I have filtered my data to only include wages and public assistance income greater than -1. This has narrowed my number of observations down from 70,131 to 59,952. 
```{r load and attach data, message=FALSE}
person_data <-read_csv('SaraArman-DataSet1_Final.csv') %>%
  filter(WAGP> -1 & PAP> -1)
```
## Continous Variables 

### My first continous variable is Wages over the past 12 months (WAGP) so I wil find the:
1. mean
2. standard deviation 
3. 95% CI for Population Mean 
4. Inter Quartile Range 
5. And then I will create a histogram.

```{r quick summary statistics}
WAGP_summary <- summary(person_data$WAGP)
```
 
### Mean
 
```{r}
WAGP_summary["Mean"]
```

### Standard Deviation

```{r}
sd(person_data$WAGP)
```


### 95% CI for Population mean 
```{r}
conf_int<- t.test(person_data$WAGP)
conf_int
```

### IQR (IQR = Q3-Q1 = 47000)

```{r}
quantile(person_data$WAGP)
```


### Histogram 

This histogram was not surprising, as it shows that most people make less than 100,000 each year, with 25,000 people making 0 a year. 
```{r}
options(scipen = 999)
ggplot(person_data, aes(x = WAGP)) +
  geom_histogram(color = "black" , bins = 75)
scale_x_continuous(name = "Wages and Income Over the Past 12 Months")
                     breaks = breaks <- seq(0, 10, by = 1)
```

### My second continous variable is Public Assistance Income Recieved over past 12 months, PAP

```{r quick summary statistics}
PAP_summary <- summary(person_data$PAP)
PAP_summary
```

### Mean
 
```{r}
PAP_summary["Mean"]
```

### Standard Deviation

```{r}
sd(person_data$PAP)
```
### 95% CI 

```{r}
conf_int <- t.test(person_data$PAP)
conf_int
```

### IQR (IQR = Q3-Q1)

```{r}
quantile(person_data$PAP)
```

### Histogram 

This histogram shows that most people recevied $0 in pubic assistance income, while a very small subset of the population recieved between $2000 and $4000.
```{r}
ggplot(person_data,
       aes(x = PAP)) + 
geom_histogram(aes(fill = ..count..), bins = 6, color = "black", weight = 3) +
scale_x_continuous(name = "Public Assistance Income",
                   breaks = seq(0, 8000, by = 2000)) +
scale_y_continuous(name = "Count",
                  breaks = seq(0, 60000, by = 5000)) + 
labs(title = 'Public Assistance Income Received Over the Past 12 Months')+
 theme_bw()
coord_flip()
```


### My third continous variable is age

```{r quick summary statistics}
AGEP_summary <- summary(person_data$AGEP)
AGEP_summary
```

### Mean
 
```{r}
AGEP_summary["Mean"]
```
### Standard Deviation

```{r}
sd(person_data$AGEP)
```
### 95% CI 

```{r}
conf_int <- t.test(person_data$AGEP)
conf_int
```
### Histogram 

This histogram shows that there are many people who live in Massachusetts who are between the ages of 20-25, and 55-65. 

```{r}
ggplot(person_data,
       aes(x = AGEP)) + 
geom_histogram(aes(fill = ..count..), bins = 20, color = "black", weight = 5) +
scale_x_continuous(name = "Age",
                   breaks = seq(0, 95, by = 5)) +
scale_y_continuous(name = "Count",
                  breaks = seq(0, 30000, by = 2500)) + 
labs(title = 'Age')+
 theme_bw()
coord_flip()
```


## Categorical Variables
For each categorical variable in my data set (SEX, SCHL, CIT), I will calculate the 95-percent confidence interval for the proportion of the population in each category.

### My first categorical variable is Sex

```{r}
unique(person_data$SEX_label)
```

```{r}
table(person_data$SEX_label)
```
```{r quick summary statistics}
SEX_summary <- summary(person_data$SEX)
SEX_summary
```

```{r}
table(person_data$SEX_label) / sum(table(person_data$SEX_label))
```

```{r}
mean(person_data$SEX_label == "Female")
mean(person_data$SEX_label == "Male")
```

```{r}
conf_int <- t.test(person_data$SEX_label == "Female")
conf_int
conf_int <- t.test(person_data$SEX_label == "Male")
conf_int
```

### My second categorical variable is Educational Attainment (SCHL) 
I am finding the confidence interval for each category within educational attainment.

```{r}
unique(person_data$SCHL_label)
```

```{r quick summary statistics}
SCHL_summary <- summary(person_data$SCHL)
SCHL_summary
```

```{r}
table(person_data$SCHL_label)
```

```{r}
conf_int <- t.test (person_data$SCHL == "1 or more years of college credit, no degree")
conf_int
conf_int <- t.test (person_data$SCHL == "12th grade - no diploma")
conf_int
conf_int <- t.test (person_data$SCHL == "Associate's degree")
conf_int
conf_int <- t.test (person_data$SCHL == "Bachelor's degree")
conf_int
conf_int <- t.test (person_data$SCHL == "Doctorate degree")
conf_int
conf_int <- t.test (person_data$SCHL == "GED or alternative credential")
conf_int
conf_int <- t.test (person_data$SCHL == "Grade 1")
conf_int
conf_int <- t.test (person_data$SCHL == "Grade 10")
conf_int
conf_int <- t.test (person_data$SCHL == "Grade 11")
conf_int
conf_int <- t.test (person_data$SCHL == "Grade 2")
conf_int
conf_int <- t.test (person_data$SCHL == "Grade 3")
conf_int
conf_int <- t.test (person_data$SCHL == "Grade 4")
conf_int
conf_int <- t.test (person_data$SCHL == "Grade 5")
conf_int
conf_int <- t.test (person_data$SCHL == "Grade 6")
conf_int
conf_int <- t.test (person_data$SCHL == "Grade 7")
conf_int
conf_int <- t.test (person_data$SCHL == "Grade 8")
conf_int
conf_int <- t.test (person_data$SCHL == "Grade 9")
conf_int
conf_int <- t.test (person_data$SCHL == "Kindergarten")
conf_int
conf_int <- t.test (person_data$SCHL == "Master's degree")
conf_int
conf_int <- t.test (person_data$SCHL == "No schooling completed")
conf_int
conf_int <- t.test (person_data$SCHL == "Nursery school, preschool")
conf_int
conf_int <- t.test (person_data$SCHL == "Professional degree beyond a bachelor's degree")
conf_int
conf_int <- t.test (person_data$SCHL == " Regular high school diploma")
conf_int
conf_int <- t.test (person_data$SCHL == "Some college, but less than 1 year")
conf_int
```


### My third categorical variable is Citizenship status , CIT
```{r}
unique(person_data$CIT_label)
```

```{r quick summary statistics}
CIT_summary <- summary(person_data$CIT)
CIT_summary
```

```{r}
table(person_data$CIT_label)
```

```{r}
table(person_data$CIT_label) / sum(table(person_data$CIT_label))
```

```{r}
conf_int <- t.test (person_data$CIT == "Born abroad of American parent(s)")
conf_int
conf_int <- t.test (person_data$CIT == "Born in Puerto Rico, Guam, the U.S. Virgin Islands, or the Northern Marianas ")
conf_int
conf_int <- t.test (person_data$CIT == "Born in the U.S.")
conf_int
conf_int <- t.test (person_data$CIT == "Not a citizen of the U.S. ")
conf_int
conf_int <- t.test (person_data$CIT == "U.S. citizen by naturalization ")
conf_int
```
