---
title: "Assignment 2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidycensus)
library(tidyverse)
library(ggplot2)
library(ggthemes)
```


```{r}
person_data <-read_csv('SaraArman-DataSet1_Final.csv') %>%
  filter(WAGP> -1 & PAP> -1 & SEX== 2)
```
## Continous Variables 
### My first continous variable is Wages over the past 12 months (WAGP) so I wil find the 

1. mean
2. standard deviation 
3. 95% CI for Population Mean 
4. Inter Quartile Range 
5. and I will create a histogram.
```{r quick summary statistics}
WAGP_summary <- summary(person_data$WAGP)
WAGP_summary
```
 
 ### Mean
 
```{r}
WAGP_summary["Mean"]
```

### Standard Deviation

```{r}
sd(person_data$WAGP)
```


### 95% CI for Population mean (if t-test is the difference of means between 2 groups, how does this work?)
```{r}
conf_int <- t.test(person_data$WAGP)
conf_int
```

### IQR (IQR = Q3-Q1?)

```{r}
quantile(person_data$WAGP)
```


### Histogram 
```{r}
options(scipen = 999)
ggplot(person_data, aes(x = WAGP)) +
  geom_histogram()
```

### My second continous variable is Public Assistance Income Recieved over past 12 months, PAP

```{r quick summary statistics}
PAP_summary <- summary(person_data$PAP)
PAP_summary
```

### Mean
 
```{r}
PAP_summary["Mean"]
```

### Standard Deviation

```{r}
sd(person_data$PAP)
```
### 95% CI 

```{r}
conf_int <- t.test(person_data$PAP)
conf_int
```

### IQR (IQR = Q3-Q1?)

```{r}
quantile(person_data$PAP)
```

### Histogram 
```{r}
ggplot(person_data,
       aes(x = PAP)) + 
geom_histogram(aes(fill = ..count..), bins = 20000, color = "black", weight = 5) +
scale_x_continuous(name = "Public Assistance Income",
                   breaks = seq(1000, 5000, by = 1000)) +
scale_y_continuous(name = "Count",
                  breaks = seq(0, 50, by = 10)) + 
labs(title = 'Public Assistance Income Received Over the Past 12 Months')+
 theme_bw()
coord_flip()
```


### My third continous variable is age

```{r quick summary statistics}
AGEP_summary <- summary(person_data$AGEP)
AGEP_summary
```

### Mean
 
```{r}
AGEP_summary["Mean"]
```
### Standard Deviation

```{r}
sd(person_data$AGEP)
```
### 95% CI 

```{r}
conf_int <- t.test(person_data$AGEP)
conf_int
```
### Histogram 
```{r}
ggplot(person_data, aes(x = AGEP)) +
  geom_histogram()
```
## Categorical Variables

### My first categorical variable is Sex

```{r}
unique(person_data$SEX_label)
```
```{r}
table(person_data$SEX_label)
```
```{r}
table(person_data$SEX_label) / sum(table(person_data$SEX_label))
```
```{r}
mean(person_data$SEX_label == "Female")
```

```{r}
t.test(person_data$SEX_label == "Female")
```
```{r}
mean(person_data$SEX_label == "Male")
```

```{r}
t.test(person_data$SEX_label == "Female")
```

### My second categorical variable is Educational Attainment SCHL 

```{r}
unique(person_data$SCHL_label)
```

```{r}
table(person_data$SCHL_label)
```

```{r}
table(person_data$SCHL_label) / sum(table(person_data$SCHL_label))
```

### Do I need to do this for every category? *Ask Carole
```{r}
mean(person_data$SCHL_label == "Regular high school diploma")
```


### My third categorical variable is Citizenship status , CIT
```{r}
unique(person_data$CIT_label)
```
```{r}
table(person_data$CIT_label)
```

```{r}
table(person_data$CIT_label) / sum(table(person_data$CIT_label))
```

```{r}
t.test(person_data$CIT_label == "Not a citizen of the U.S.")
```